Percussive Synthesizer Generation with Virtual Listeners

### 2020-10-09 templates_aimusic2020/paper.tex
appendix shortened

\documentclass[runningheads,a4paper]{llncs}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{apacite}
%added 
\usepackage[inline]{enumitem}    
\usepackage{subcaption}
\usepackage{caption}
\usepackage{threeparttable} %annotations for table
\usepackage{csquotes}
\usepackage{appendix}
\usepackage{comment}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\usepackage[bottom]{footmisc}
\usepackage{makecell}
\usepackage{xcolor}

@@

In order to classify sounds, we need to summarize them as features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT)~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used for its reconstruction, perhaps it can be utilized as a source of fundamental features~\cite{lee2009unsupervised,hershey2017cnn}. We defined 3 STFT (or FFT) transformation functions which we believe to capture important, unique attributes of percussive sounds. 
\begin{enumerate}
\item Envelope Transformation: Represents changes in loudness for the duration of the signal. Using STFT we generate a matrix $M_{i \times j}$ with rows $i$ and columns $j$ corresponding to time steps and frequency bins. Values $v_{i \times j}$ indicate the magnitude of frequency bin $j$ at each time-step $i$. We approximate the envelope of the signal by summing the values of $M$ at each time-step. 
\item Frequency Transformation:  Similar to envelope calculation, but the summation is done along the frequency axis. Shorter hop-sizes and wider windows were used to increase frequency resolution. 
\item Spectrum Transformation: Mel Scaled STFT with values normalized from 0-1. 
\end{enumerate}

@@

\bibliographystyle{apacite}
\bibliography{citations.bib}

\begin{appendices}
%======================
\chapter{Datasets, Details, and Downloads}
\label{appendix:datasets}
\subsection{Download Instructions}
FreeDB, survey data, and randomly generated sounds can be downloaded from: \url{https://zenodo.org/record/3994999}\\
RadarDB requires lengthy downloads and processing. The script for its automated creation can be found under the \enquote{getting\_data} directory of our project code: \footnote{\url{https://github.com/imilas/Synths_Stacks_Search}}
\subsection{Details}
Our data sources are a large set of 2 second or shorter drum samples aggregated from personal libraries, free drum kits from the sample-swap project \footnote{https://sampleswap.org/} which we organized into 5 groups, and a large set of drum sounds aggregated from royalty free sources such as music radar \footnote{https://www.musicradar.com/}. We put together 3 databases of drums using these sources. We also created a database of synthetic noise from 1, 3, and 5 stacked virtual synthesizers. Specifications about our datasets can be found in tables~\ref{table:all_db}. We have made our dataset of free-drum sounds available for download. The scripts used to download and process royalty free samples are also made available.

@@

%==========================
\chapter{Audio Transformation Functions}
\label{appendix:trans_funcitons}
Figure ~\ref{stackspectrums} contains the graphed representation of features extracted for 3 different samples. Sample $a$ is a recorded hat from our database. sample $b$ is an example of randomly generated noise with percussive qualities that we found suitably similar to a snare sound. Sample $c$ is an example of a randomly generated noise where the spectrum features are necessary for proper classification.
\begin{figure}[h]
\centering
\textbf{Visual Representation of Raw Features}\par\medskip
    \subcaptionbox{Recorded hat sample}{    \includegraphics[width=1\columnwidth]{images/ff1.pdf}
    }
    \subcaptionbox{Randomly generated audio with percussive qualities, resembling a tight snare}{\includegraphics[width=1\columnwidth]{images/ff2.pdf}}
    \subcaptionbox{A randomly generated noise with a percussive envelop but non-percussive frequency features (modulated pitch)}
    { \includegraphics[width=1\columnwidth]{images/ff3.pdf}}
\label{stackspectrums}
\end{figure}
% \begin{enumerate}
% \item Envelope Transformation: Represents changes in loudness for the duration of the signal. Using STFT we generate a matrix $M_{i \times j}$ with rows $i$ and columns $j$ corresponding to time steps and frequency bins respectively. Values $v_{i \times j}$ indicating the magnitude of the frequency bin $j$ at each time-step $i$. Information about the envelope of the signal can be extracted by summing the values of $M$ for each time-step (or row $i$), giving us a feature vector that is normalized to the range of 0 to 1. The information contained in this vector is an alternative to Root-Mean-Square measurements of a sliding window over the signal.
% \item Frequency Transformation: A static, normalized snap-shot of the the frequencies present within the audio. The calculation of this feature vector is similar to the envelope, but the summation is done along the frequency axis. Another important distinction is that since capturing an adequate frequency resolution is important for this transformation, we utilized shorter hop-sizes and wider windows. A Mel Scale transformation was also applied in hopes that the captured features better represent human perception of frequencies. 
% \item Spectrum Transformation: This function is simply a Mel Scaled STFT with its values normalized from 0-1. Spectrograms contain more information about the original signal but their analysis requires more complex computational methods.
% \end{enumerate}

@@

\begin{figure}[htb!]
    \begin{center}
    \textbf{Cross Validation F-Scores For All Sound Groups}\par\medskip
    \makebox[\textwidth]{\includegraphics[width=0.8\paperwidth]{images/mme_comparisons_mme.pdf}}
    \caption{Boxplots visualizing the F-Score results for each cross-validation. The individual scores, means, medians, standard-deviation and outliers are depicted. The differences are noticeable, yet means lie within the \%88-92 range. Envelope features improve classification accuracy for all models. }
    \label{fig:f1_allg_box}
    \end{center}
\end{figure}

We're also interested in how these models perform on the binary drum vs not drum task.
After grouping all drums together, we repeat the model selection process above. We also repeat the hyper-parameter optimization step where no changes appeared necessary except for a reduction in the number of neighbors for the K-Neighbor model (from 30 to 5). As show in in Figures~\ref{fig:f1_allg_box} and~\ref{fig:f1_dvn_box}, the addition of envelope features had a positive effect on performance for all mdoels, yet the RandomForest and ExtraTrees models clearly outperform the other classifiers in both tasks. We train the top two models on \%80 of our database and use the remaining \%20 to create the confusion matrices and the F-Scores shown in Figures~\ref{fig:conf_f1_dvn}.
\begin{figure}[h!]   
    \begin{center}
        \textbf{Cross Validation F-Scores For Drum Vs Not-Drum}
    \makebox[\textwidth]{\includegraphics[width=0.8\paperwidth]{images/mme_comparisons_dvn.pdf}}
    \caption{F-Score results for each cross-validation. Models perform better as there are less categorization groups. Envelope features increase accuracy for all models. Random Forest and Extra Trees remain the top two models. }
    \label{fig:f1_dvn_box}
    \end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
    \textbf{ Classification Report for DvDvN and DvN  }\par\medskip
    \makebox[\textwidth]{
    \subfloat[Precision, recall, F1-Score, and number of supporting examples]{ \includegraphics[width=9cm,height=9cm]{images/f1_mme.pdf}
    \includegraphics[width=9cm,height=9cm]{images/f1_dvn.pdf}
    }

@@

\begin{center}
    \makebox[\textwidth]{
    \subfloat[Confusion matrices]{ \includegraphics[width=9cm,height=9cm]{images/conf_mme.pdf}
    \includegraphics[width=9cm,height=9cm]{images/conf_dvn.pdf}
    }}
\end{center}

\caption{F-Scores and confusion matrix of ExtraTrees model for both DvDvN and drum vs not-drum categorization.}
\label{fig:conf_f1_dvn}
\end{figure}

Based on these reports, having multiple options for drum categorization does not noticeably influence our models accuracy in categorizing synthetic noise as synthetic noise. However, the DvDvN models's slightly smaller false negative rate for synthetic noise (11 vs 13 false negatives) is countered by a slightly higher rate of categorizing drums as synthetic noise (12 vs 9).  We therefore use the DvDvN implementation as it simultaneously categorizes drum types and separates noise from drums. However, without manual inspection, we cannot confirm the extent of this model's usefulness. 
%==============
\chapter{AutoEncoder Structures and Hyper-Parameter Optimization}
\label{appendix:hyperparam}
% could do multi-objective pareto front stuff too
Within the context of machine learning, a model's \emph{hyper-parameters} are fixed parameters which are set before the training begins (e.g number of layers, size of layers, loss function) and are not learned during the minimization procedure~\cite{bengio2000gradient}. To assist us with the construction of our model, we defined a number of possible choices for the architecture of our model and audio-transformers and used hyper-parameter optimization to extract promising sets of values~\ref{table:hyper_params}.

@@

Using the optuna optimization tools \cite{akiba2019optuna}, we conducted 500 search trials. The trial's success is measured in their final loss value, calculated by applying the model to test data-set. Each trials consisted of 20 epoch of training using a selected set of hyper-parameters. An additional 100 trials with 40 epochs of training were conducted following the initial 500 to test the effect of longer epochs. Each trial's intermediate results (loss at every n epoch,s where $0<n<20$) were reported to a multi-armed bandit based pruner for early stoppage of unpromising trials~\cite{li2017hyperband}. We employed a tree-structured parzen estimator for better navigation of the search space~\cite{bergstra2011algorithms,akiba2019optuna} but found short reversions to a random sampling coupled with a decrease in the frequency of pruning helpful in exiting local minima. \\
\begin{table}[htb]
\begin{tabular}{|p{28mm}|p{50mm}|p{21mm}|p{21mm}|}
\hline
Hyper-Param. & Description  & Values & Distribution\\ \hline
Model Type      &   Affects encoder's first hidden layer & CNN,FC & Categorical \\  \hline
Optimizer       & Updates network's weights based on loss & Adam,SGD & Categorical  \\  \hline
Hidden Layers   & Extra hidden layer for the Encoder & True,False & Categorical \\  \hline
Time Steps & Temporal granularity of the spectrogram. Affects FFT windowing. & 10,20 & Categorical  \\ \hline
Learning Rate   &    Optimizer's learning rate  & $1^{-4}$ ... $1^{-1}$ & Uniform      \\ \hline
Frequency Bins & Number of spectrogram frequency bins & 10, 30,60 & Categorical \\ \hline

@@

\begin{figure}[htb]
\centering
\textbf{Tracing the Best Loss Values}
\includegraphics[width=12cm,height=7cm]{images/Optimization_History.pdf}
\caption{Best loss values found during hyper-parameter optimization. The effect of a switch to random sampling and an increase of the pruning threshold can be observed during trials 270 and 310.}
\label{chap3:bestvalues}
\end{figure}

\begin{figure}
\begin{center}
    \textbf{Hyper-Parameters' Loss Correlation}
    \makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{images/slice_plot.pdf}}
      \caption{Sliced plot depicting the correlation between hyper-parameters and loss values. The color-scale shows the number of times each parameters has been used in a trial. Our sampling algorithm aims to utilize spaces with higher potential more often. }
    \label{fig:slicegraph}
    \end{center}
\end{figure}

\begin{table}[htb!]
\centering
\begin{tabular}{|p{6cm}|p{6cm}|}
\hline
Hyper-Param. & Value  \\ \hline
Model Type      &  CNN  \\ \hline
Optimizer       & Adam  \\ \hline
Hidden Layers   & Not Applicable  \\\hline
Learning Rate   &  0.001145\\ \hline
Frequency Bins & 30 \\ \hline
Time Steps & 20 \\ \hline
Latent Size & 64 \\ \hline
Regularization & 3.25^{-6}\\ \hline
Dropout Rate & 0.5 \\ \hline
\end{tabular}
\caption{Top performing hyper-parameter set}
\label{table:best_params}
\end{table}

>>>>

\documentclass[runningheads,a4paper]{llncs}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{apacite}
%added 
\usepackage[inline]{enumitem}    
\usepackage{subcaption}
\usepackage{caption}
\usepackage{threeparttable} %annotations for table
\usepackage{csquotes}
\usepackage[title]{appendix}
\usepackage{comment}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\usepackage[bottom]{footmisc}
\usepackage{makecell}
\usepackage{xcolor}

@@

In order to classify sounds, we need to summarize them as features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT)~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used for its reconstruction, perhaps it can be utilized as a source of fundamental features~\cite{lee2009unsupervised,hershey2017cnn}. We defined 3 STFT (or FFT) transformation functions to capture important features of percussive sounds. 
\begin{enumerate}
\item Envelope Transformation: Represents changes in loudness for the duration of the signal. Using STFT we generate a matrix $M_{i \times j}$ with rows $i$ and columns $j$ corresponding to time steps and frequency bins. Values $v_{i \times j}$ indicate the magnitude of frequency bin $j$ at each time-step $i$. We approximate the envelope of the signal by summing the values of $M$ at each time-step. 
\item Frequency Transformation:  Similar to envelope calculation, but the summation is done along the frequency axis. Shorter hop-sizes and wider windows were used to increase frequency resolution. 
\item Spectrum Transformation: Mel Scaled STFT. Values normalized from 0-1. 
\end{enumerate}

@@

\bibliographystyle{apacite}
\bibliography{citations.bib}
\clearpage

\begin{appendices}
\markboth{Appendix}{Appendix}

\chapter{Datasets, Details, and Downloads}
\label{appendix:datasets}
\subsection{Download Instructions}
FreeDB, survey data, and randomly generated sounds can be downloaded from: \url{https://zenodo.org/record/3994999}\\
RadarDB requires lengthy downloads and processing. The script for its automated creation can be found under the \enquote{getting\_data} directory of our project code: \footnote{\url{https://github.com/imilas/Synths_Stacks_Search}}
\subsection{Details}
Our data sources are a large set of 2 second or shorter drum samples aggregated from personal libraries, free drum kits from the sample-swap project \footnote{https://sampleswap.org/} which we organized into 5 groups, and a large set of drum sounds aggregated from royalty free sources such as music radar \footnote{https://www.musicradar.com/}. We put together 3 databases of drums using these sources. We also created a database of synthetic noise from 1, 3, and 5 stacked virtual synthesizers. Specifications about our datasets can be found in tables~\ref{table:all_db}. We have made our dataset of free-drum sounds available for download. The scripts used to download and process royalty free samples are also made available.

@@

%==========================
\chapter{Audio Transformation Functions}
\label{appendix:trans_funcitons}
Figure~\ref{fig:stackspectrums} contains the graphed representation of features extracted for 3 different samples. Sample $a$ is a recorded hat from our database. sample $b$ is an example of randomly generated noise with percussive qualities that we found suitably similar to a snare sound. Sample $c$ is an example of a randomly generated noise where the spectrum features are necessary for proper classification.
\begin{figure}
\centering
\textbf{Visual Representation of Raw Features}\par\medskip
    \subcaptionbox{Recorded hat sample}{    \includegraphics[width=1\columnwidth]{images/ff1.pdf}
    }
    \subcaptionbox{Randomly generated audio with percussive qualities, resembling a tight snare}{\includegraphics[width=1\columnwidth]{images/ff2.pdf}}
    \subcaptionbox{A randomly generated noise with a percussive envelop but non-percussive frequency features (modulated pitch)}
    { \includegraphics[width=1\columnwidth]{images/ff3.pdf}}
\caption{Graphed representation of features extracted for 3 different samples. }
\label{fig:stackspectrums}
\end{figure}

% \begin{enumerate}
% \item Envelope Transformation: Represents changes in loudness for the duration of the signal. Using STFT we generate a matrix $M_{i \times j}$ with rows $i$ and columns $j$ corresponding to time steps and frequency bins respectively. Values $v_{i \times j}$ indicating the magnitude of the frequency bin $j$ at each time-step $i$. Information about the envelope of the signal can be extracted by summing the values of $M$ for each time-step (or row $i$), giving us a feature vector that is normalized to the range of 0 to 1. The information contained in this vector is an alternative to Root-Mean-Square measurements of a sliding window over the signal.
% \item Frequency Transformation: A static, normalized snap-shot of the the frequencies present within the audio. The calculation of this feature vector is similar to the envelope, but the summation is done along the frequency axis. Another important distinction is that since capturing an adequate frequency resolution is important for this transformation, we utilized shorter hop-sizes and wider windows. A Mel Scale transformation was also applied in hopes that the captured features better represent human perception of frequencies. 
% \item Spectrum Transformation: This function is simply a Mel Scaled STFT with its values normalized from 0-1. Spectrograms contain more information about the original signal but their analysis requires more complex computational methods.
% \end{enumerate}

@@

\begin{figure}[htb!]
\begin{center}
    \textbf{Cross Validation F-Scores For All Sound Groups}\par\medskip
    \makebox[\textwidth]{\includegraphics[width=0.7\paperwidth]{images/mme_comparisons_mme.pdf}}
    \caption{Boxplots visualizing the F-Score results for each cross-validation. The individual scores, means, medians, standard-deviation and outliers are depicted. Envelope features improve classification accuracy. }
    \label{fig:f1_allg_box}
\end{center}

\begin{center}
        \textbf{Cross Validation F-Scores For Drum Vs Not-Drum}
    \makebox[\textwidth]{\includegraphics[width=0.7\paperwidth]{images/mme_comparisons_dvn.pdf}}
    \caption{F-Score results for each cross-validation. Models perform better as there are less categorization groups.  }
    \label{fig:f1_dvn_box}
\end{center}

\label{fig:conf_f1_dvn}
\end{figure}

% \begin{figure}[h!]
%     \begin{center}
%     \textbf{Cross Validation F-Scores For All Sound Groups}\par\medskip
%     \makebox[\textwidth]{\includegraphics[width=0.8\paperwidth]{images/mme_comparisons_mme.pdf}}
%     \caption{Boxplots visualizing the F-Score results for each cross-validation. The individual scores, means, medians, standard-deviation and outliers are depicted. The differences are noticeable, yet means lie within the \%88-92 range. Envelope features improve classification accuracy for all models. }
%     \label{fig:f1_allg_box}
%     \end{center}
% \end{figure}
% \begin{figure}[h!]   
%     \begin{center}
%         \textbf{Cross Validation F-Scores For Drum Vs Not-Drum}
%     \makebox[\textwidth]{\includegraphics[width=0.8\paperwidth]{images/mme_comparisons_dvn.pdf}}
%     \caption{F-Score results for each cross-validation. Models perform better as there are less categorization groups. Envelope features increase accuracy for all models. Random Forest and Extra Trees remain the top two models. }
%     \label{fig:f1_dvn_box}
%     \end{center}
% \end{figure}
% We're also interested in how these models perform on the binary drum vs not drum task.
% After grouping all drums together, we repeat the model selection process above. We also repeat the hyper-parameter optimization step where no changes appeared necessary except for a reduction in the number of neighbors for the K-Neighbor model (from 30 to 5). As show in in Figures~\ref{fig:f1_allg_box} and~\ref{fig:f1_dvn_box}, the addition of envelope features had a positive effect on performance for all mdoels, yet the RandomForest and ExtraTrees models clearly outperform the other classifiers in both tasks. We train the top two models on \%80 of our database and use the remaining \%20 to create the confusion matrices and the F-Scores shown in Figures~\ref{fig:conf_f1_dvn}.

\begin{figure}[htb!]
\begin{center}
    \textbf{ Classification Report for DvDvN and DvN  }\par\medskip
    \makebox[\textwidth]{
    \subfloat[Precision, recall, F1-Score, and number of supporting examples]{ \includegraphics[width=7cm,height=7cm]{images/f1_mme.pdf}
    \includegraphics[width=7cm,height=7cm]{images/f1_dvn.pdf}
    }

@@

\begin{center}
    \makebox[\textwidth]{
    \subfloat[Confusion matrices]{ \includegraphics[width=7cm,height=7cm]{images/conf_mme.pdf}
    \includegraphics[width=7cm,height=7cm]{images/conf_dvn.pdf}
    }}
\end{center}
\caption{F-Scores and confusion matrix of ExtraTrees model for both DvDvN and DvN.}
\label{fig:conf_f1_dvn}
\end{figure}

% Based on these reports, having multiple options for drum categorization does not noticeably influence our models accuracy in categorizing synthetic noise as synthetic noise. However, the DvDvN models's slightly smaller false negative rate for synthetic noise (11 vs 13 false negatives) is countered by a slightly higher rate of categorizing drums as synthetic noise (12 vs 9).  We therefore use the DvDvN implementation as it simultaneously categorizes drum types and separates noise from drums. However, without manual inspection, we cannot confirm the extent of this model's usefulness. 
%==============
\chapter{AutoEncoder Structures and Hyper-Parameter Optimization}
\label{appendix:hyperparam}
% could do multi-objective pareto front stuff too
Within the context of machine learning, a model's \emph{hyper-parameters} are fixed parameters which are set before the training begins (e.g number of layers, size of layers, loss function) and are not learned during the minimization procedure~\cite{bengio2000gradient}. To assist us with the construction of our model, we defined a number of possible choices for the architecture of our model and audio-transformers and used hyper-parameter optimization to extract promising sets of values~\ref{table:hyper_params}.

@@

Using the optuna optimization tools \cite{akiba2019optuna}, we conducted 500 search trials. The trial's success is measured in their final loss value, calculated by applying the model to test data-set. Each trial consisted of 20 epochs of training.

\begin{table}[b!]
\centering
\begin{tabular}{|p{6cm}|p{6cm}|}
\hline
Hyper-Param. & Value  \\ \hline
Model Type      &  CNN  \\ \hline
Optimizer       & Adam  \\ \hline
Hidden Layers   & Not Applicable  \\\hline
Learning Rate   &  0.001145\\ \hline
Frequency Bins & 30 \\ \hline
Time Steps & 20 \\ \hline
Latent Size & 64 \\ \hline
Regularization & 3.25^{-6}\\ \hline
Dropout Rate & 0.5 \\ \hline
\end{tabular}
\caption{Top performing hyper-parameter set}
\label{table:best_params}
\end{table}
\begin{table}[htb]
\begin{tabular}{|p{28mm}|p{50mm}|p{21mm}|p{21mm}|}
\hline
Hyper-Param. & Description  & Values & Distribution\\ \hline
Model Type      &   Affects encoder's first hidden layer & CNN,FC & Categorical \\  \hline
Optimizer       & Updates network's weights based on loss & Adam,SGD & Categorical  \\  \hline
Hidden Layers   & Extra hidden layer for the Encoder & True,False & Categorical \\  \hline
Time Steps & Temporal granularity of the spectrogram. Affects FFT windowing. & 10,20 & Categorical  \\ \hline
Learning Rate   &    Optimizer's learning rate  & $1^{-4}$ ... $1^{-1}$ & Uniform      \\ \hline
Frequency Bins & Number of spectrogram frequency bins & 10, 30,60 & Categorical \\ \hline

@@

% \begin{figure}[htb]
% \centering
% \textbf{Tracing the Best Loss Values}
% \includegraphics[width=12cm,height=7cm]{images/Optimization_History.pdf}
% \caption{Best loss values found during hyper-parameter optimization. The effect of a switch to random sampling and an increase of the pruning threshold can be observed during trials 270 and 310.}
% \label{chap3:bestvalues}
% \end{figure}

\begin{figure}[tb]
\begin{center}
    \textbf{Hyper-Parameters' Loss Correlation}
    \makebox[\textwidth]{\includegraphics[width=0.95\paperwidth]{images/slice_plot.pdf}}
      \caption{Sliced plot depicting the correlation between hyper-parameters and loss values. The color-scale shows the number of times each parameters has been used in a trial. Our sampling algorithm aims to utilize spaces with higher potential more often. }
    \label{fig:slicegraph}
    \end{center}
\end{figure}
